% Quantitative Schedule Risk Analysis (QSRA) — Methodology
% Compile: pdflatex methodology.tex
\documentclass[10pt]{article}

\usepackage[top=0.7in,bottom=0.8in,left=0.75in,right=0.75in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{libertine}
\usepackage[scaled=0.85]{beramono}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage{parskip}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{xcolor}

\definecolor{muted}{HTML}{6B6B6B}
\definecolor{text}{HTML}{1A1A1A}

\titleformat{\section}
  {\normalfont\ttfamily\footnotesize\bfseries\color{muted}\MakeUppercase}
  {}{0pt}{}
\titlespacing*{\section}{0pt}{4pt}{0pt}

\setlist[enumerate]{leftmargin=*,itemsep=0pt,topsep=2pt,parsep=0pt}
\setlist[itemize]{leftmargin=*,itemsep=0pt,topsep=2pt,parsep=0pt}

\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0.4pt}
\fancyfoot[L]{\ttfamily\scriptsize\color{muted}DOC-QSRA-METH-01}
\fancyfoot[C]{\ttfamily\scriptsize\color{muted}REV 2026-04-10}
\fancyfoot[R]{\ttfamily\scriptsize\color{muted}PAGE \thepage}

\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{3pt}
\setlength{\abovedisplayshortskip}{1pt}
\setlength{\belowdisplayshortskip}{1pt}
\setlength{\parskip}{2pt}

\begin{document}

\noindent
{\ttfamily\small\bfseries QUANTITATIVE SCHEDULE RISK ANALYSIS (QSRA) \textbar{} METHODOLOGY}\\[2pt]
{\ttfamily\scriptsize\color{muted}CPM FORWARD PASS \textbar{} TRIANGULAR DURATION UNCERTAINTY \textbar{} MONTE CARLO SAMPLING \textbar{} SPEARMAN SENSITIVITY \textbar{} AACE RP 57R-09 COMPLIANT}

\vspace{4pt}
\hrule height 0.4pt
\vspace{4pt}

\section{Setup}
The project is modelled as a Work Breakdown Structure (WBS)-organised directed acyclic graph of $n$ activities with finish-to-start (FS), zero-lag precedence relationships. For each activity $i$, $P(i)$ denotes its set of immediate predecessors and $d_i > 0$ its baseline duration in working days. The methodology follows AACE International Recommended Practice 57R-09 (\emph{Integrated Cost and Schedule Risk Analysis}) and replicates the core iteration engine of Oracle Primavera Risk Analysis (OPRA).

\section{Deterministic Baseline (CPM Forward Pass)}
Given a topological ordering of the activity network, the Critical Path Method (CPM) forward pass computes the early start ($ES_i$) and early finish ($EF_i$) of each activity recursively:
\begin{equation}
ES_i = \begin{cases} 0 & P(i) = \emptyset \\ \displaystyle\max_{j \in P(i)} EF_j & \text{otherwise} \end{cases}, \qquad EF_i = ES_i + d_i, \qquad T^{\text{det}} = \max_{i} EF_i.
\end{equation}
$T^{\text{det}}$ is the deterministic project finish under the assumption that every activity executes at exactly its baseline duration.

\section{Duration Uncertainty Model}
Each activity receives a \emph{three-point estimate} consisting of multiplicative optimistic, most-likely, and pessimistic factors $(\alpha_i, \mu_i, \beta_i)$ with $0 < \alpha_i \le \mu_i \le \beta_i$, yielding
\begin{equation}
(a_i, m_i, b_i) = (\alpha_i d_i,\, \mu_i d_i,\, \beta_i d_i).
\end{equation}
Activity duration is modelled as a triangular random variable $D_i \sim \mathrm{Tri}(a_i, m_i, b_i)$ with density
\begin{equation}
f_{D_i}(x) =
\begin{cases}
\dfrac{2(x - a_i)}{(b_i - a_i)(m_i - a_i)} & a_i \le x \le m_i, \\[0.2em]
\dfrac{2(b_i - x)}{(b_i - a_i)(b_i - m_i)} & m_i < x \le b_i, \\[0.2em]
0 & \text{otherwise.}
\end{cases}
\end{equation}
Risk-register findings are encoded into distribution \emph{width}: activities exposed to multiple risks receive wider pessimistic tails ($\beta_i$). Consistent with AACE RP 57R-09, \emph{discrete risk events} (explicit probability--impact pairs) are not modelled; all risk is represented through duration uncertainty, trading event-level fidelity for parameter transparency and auditability.

\section{Monte Carlo Procedure}
For $N$ iterations (default $N = 10{,}000$, fixed pseudo-random seed for reproducibility), each iteration $k$ samples independent durations $d_i^{(k)} \sim \mathrm{Tri}(a_i, m_i, b_i)$ for every activity, executes the CPM forward pass to obtain $EF_i^{(k)}$, and records the project finish $T^{(k)} = \max_{i} EF_i^{(k)}$.

\section{Output Statistics and Risk Exposure}
The empirical distribution $\{T^{(k)}\}_{k=1}^{N}$ yields
\begin{equation}
\bar{T} = \frac{1}{N}\sum_{k=1}^{N} T^{(k)}, \quad
s_T^2 = \frac{1}{N}\sum_{k=1}^{N}(T^{(k)} - \bar{T})^2, \quad
\hat{F}_T(t) = \frac{1}{N}\sum_{k=1}^{N} \mathbf{1}\{T^{(k)} \le t\}.
\end{equation}
Confidence levels are reported as percentiles $P_p = \hat{F}_T^{-1}(p/100)$ for $p \in \{5, 10, \ldots, 95\}$ and visualised on the histogram and the S-curve (empirical CDF). Two scalars carry primary operational weight: the \textbf{deterministic confidence} $\hat{F}_T(T^{\text{det}})$ --- a low value ($\ll 50\%$) indicates the baseline is unreachable --- and the \textbf{risk exposure at P80}, $\Delta_{80} = P_{80} - T^{\text{det}}$, which represents the cumulative schedule buffer required to achieve 80\% delivery confidence and serves as the basis for commitment-date negotiation.

\section{Activity Sensitivity (Tornado)}
For each activity $i$, the driver effect on project finish is quantified by the Spearman rank correlation between its sampled duration and the iteration finish:
\begin{equation}
\rho_i = 1 - \frac{6 \sum_{k=1}^{N}\bigl(R(d_i^{(k)}) - R(T^{(k)})\bigr)^2}{N(N^2 - 1)},
\end{equation}
where $R(\cdot)$ denotes the rank within the $N$-sample. Spearman is preferred over Pearson for robustness to monotone nonlinearity and heavy-tailed distributions. Activities are ranked by $|\rho_i|$ and displayed as a tornado diagram. High-$|\rho_i|$ activities need not lie on the deterministic critical path: iteration exposes \emph{near-critical path} activities---those with low deterministic total float that become critical under specific duration realisations---which is the principal motivation for QSRA over deterministic CPM alone.

\end{document}
